\documentclass{book}

\title{wordpress}
\usepackage[utf8]{inputenc}
\usepackage[margin=1.25in]{geometry}
\usepackage[colorlinks]{hyperref}
\usepackage{todonotes}
\usepackage{listings}
\lstset{basicstyle=\ttfamily,
  showstringspaces=false,
  commentstyle=\color{red},
  keywordstyle=\color{blue}
}
\usepackage{}

\author{Shady Ganem}


\def \bashversion {bash 1.2}

\begin{document}
\maketitle
{
\hypersetup{linkcolor=black}
\tableofcontents
}
\part{}
\chapter{secure the server}
The subject of security is of major importance when it comes to web development.
when deploying a new website we should assume that every transaction between the 
server and the client is compromised and must be secured. 

\section{Auto udpate}
security issues can be caused due to out dated software. we will install a tool for 
automating the process
In ssh terminal
\begin{lstlisting}[language=bash]
$ sudo apt update
$ sudo apt dist-upgrade
\end{lstlisting}
Now we will install the unattneded upgrade tool
\begin{lstlisting}[language=bash]
$ sudo dpkg-reconfigure --priority=low unattended-upgrades
\end{lstlisting}
Hit "YES"

\section{Create a user}
If you are running as root you should create a user in the sudo group.
it is considered bad practice to operate as root in linux. 

\begin{lstlisting}[language=bash]
$ adduser \username
\end{lstlisting}
Add the user to the sudo group
\begin{lstlisting}[language=bash]
$ \username -aG sudo \username
\end{lstlisting}

\section{Authentication key pair}
Logging in to the ssh server with a password is not a bad practice but it is less
secure than logging in using and authentication keys.

first lets create a directory that will hold they keys for our user.
\begin{lstlisting}[language=bash]
$ mkdir ~/.ssh && chmod 700 ~/.ssh
\end{lstlisting}

\chapter{Themes}
\section{Introduction}
\section{Hello World Theme}
In order to make sure that every thing is working we will start by creating a "Hello Wrold!" Theme.\\
All themes are stored at a "path\_to\_wp/wp-content/themes"\\
WordPress themes are made of "Template Files". Tempalte files are PHP files that contain HTML and PHP code, which 
define the look and design of the website. When someone visits the website, Wordpress searchs for a template file
under the theme folder. According To the request WordPress will parse the PHP code and return an HTML.//
The most critical template file is "index.php".



\part{}
\end{document}
